Impact of engines scheduler on the performance of Big Data pipelines in HPC

### 2021-01-15 latex/paper.tex
Partial draft for the paper.

\title{A performance comparison of Dask and Apache Spark for data-intensive neuroimaging pipelines}

@@

\begin{abstract}

In the past few years, neuroimaging has entered the Big Data era due to the joint
increase in image resolution, data sharing, and study sizes. However, no particular
Big Data engines have emerged in this field, and several alternatives remain
available. We compare two popular Big Data engines with Python APIs, Apache Spark and
Dask, for their runtime performance in processing neuroimaging pipelines. Our
evaluation uses two synthetic pipelines processing the \SI{81}{\giga\byte} BigBrain
image, and a real pipeline processing anatomical data from more than 1,000 subjects.
We benchmark these pipelines using various combinations of task durations, data
sizes, and numbers of workers, deployed on an 8-node (8 cores ea.) compute cluster in
Compute Canada's Arbutus cloud. We evaluate PySpark's RDD API against Dask's Bag,
Delayed and Futures. Results show that despite slight differences between Spark and
Dask, both engines perform comparably. However, Dask pipelines risk being limited
by Python's GIL depending on task type and cluster configuration. In all cases, the
major limiting factor was data transfer. While either engine is suitable for
neuroimaging pipelines, more effort needs to be placed in reducing data transfer time.
\end{abstract}

\begin{IEEEkeywords}
Dask, Apache Spark, performance, neuroimaging
\end{IEEEkeywords}

\section{Introduction}
% Context
The recent rise in data sharing and improved data collection strategies have brought
neuroimaging to the Big Data era~\cite{ALFAROALMAGRO:18, van2014human}. Existing
neuroimaging workflow engines, such as Nipype~\cite{Nipype:11}, are well suited for
processing the standard compute-intensive neuroimaging pipelines, but lack
incorporated Big Data strategies (i.e. in-memory computing, data locality, and
lazy-evaluation) to improve performance of the increasingly prevalent data-intensive
pipelines. As was noted in~\cite{hayot2019performance}, in-memory computing, coupled
with data locality, can bring significant performance improvements to data intensive
neuroimaging pipelines. In this work, we investigate how performance benchmarks
generalize across Big Data engines. In particular, we study the differences between
Dask~\cite{Dask:15} and Apache Spark~\cite{Spark:16}, for their suitability in the
processing of neuroimaging pipelines. Our goal is to test whether Spark or Dask has a
clear performance advantage to process Big neuroimaging Data.

%Similarities
Spark and Dask both offer in-memory computing, data locality, and lazy evaluation,
which is common for Big Data engines. Both their schedulers operate dynamically,
which is good when task runtimes are not known ahead of time~\cite{Dask:15}. They
also provide rich, high-level programming APIs, and support a variety of
infrastructure schedulers, such as Mesos~\cite{hindman2011mesos}, YARN~\cite{vavilapalli2013apache}, or HPC clusters (Dask only). Over
these similarities, the engines have differences.

First and foremost, Spark is written in Scala while Dask is in Python.
Given the popularity of Python in scientific communities, this arguably
gives an edge to Dask due to data serialization costs from Python to Scala.
On the other hand, Python's Global Interpreter Lock (GIL) might reduce
parallelism in some cases. This difference in programming languages also
has qualitative implications. As part of the \href{http://scipy.org}{SciPy}
ecosystem, Dask provides almost transparent parallelization of applications
manipulating NumPy arrays or Pandas data frames. On the other hand, Spark's
Java, R and Python APIs allow to easily parallelize analyses that combine
these languages, with reduced performance loss. Our study focuses on
performance, although we recognize that other factors will also impact the
choice of an engine in practice.

Spark and Dask were both included in the evaluation reported
in~\cite{Mehta:17}, where a neuroimaging application
processed approximately \SI{100}{\giga\byte} of data. In this work, Dask was
reported to have a slight performance advantage over Spark. Overall, Dask's
end-to-end time (makespan) was measured to be up to 14\% faster than Spark,
due to ``more efficient pipelining'' %(Fig 10) 
and serialization time to
Python. % (Fig. 12-a)
Dask, however, was reported to have a larger startup
time than Spark. % (Fig. 12-c)
The analysis remained at a quite high level
though, leaving most of the observed performance difference unexplained. In
comparison, our study will provide a detailed analysis of performance
differences and similarities between Spark and Dask for neuroimaging data processing.

The next section details the design of our benchmarking experiments. We consider two
data-intensive neuroimaging applications: high-resolution imaging,
represented by the BigBrain data~\cite{Amunts:13}, and large functional
MRI studies, represented by data from the consortium for reliability and
reproducibility (CoRR~\cite{zuo2014open}). We test application pipelines
involving different patterns (map-only, map-reduce), and different types of
implementations (plain Python, command-line,
containerized). We evaluate performance on a dedicated cluster
representative of the ones used in today's data-intensive neuroimaging
studies, using the main Dask and Spark APIs. The other sections present
our results, discussion, and conclusions.

%%%%% MATERIAL AND METHODS %%%%%%
\section{Material and Methods}

\subsection{Engines}

\subsubsection{Apache Spark} Apache Spark is a widely-used 
general-purpose Big Data engine. Its main abstraction, the Resilient Distributed
Dataset (RDD)~\cite{RDD}, is a fault-tolerant, parallel collection of data elements.
It achieves fault tolerance through the recording of data lineage, the sequence of
operations used to modify the original data. The RDD is the basis of Spark's other
data structures, namely, Datasets and DataFrames. Datasets are similar to RDDs, but
additionally use Spark SQL's optimized execution engine to further improve
performance. DataFrames, used to process tabular data, are Datasets where the data is
organized into named-columns. While the DataFrame API exists in all the available
language APIs, Datasets are limited to Scala and Java.

As data transfers in Big Data workflows are an important source of performance
bottlenecks, Spark incorporates the concepts of data locality and in-memory
computing. Data locality, popularized by MapReduce~\cite{dean2008mapreduce},
schedules tasks as close as possible to where the data is stored. In-memory computing
ensures data is maintained in memory whenever possible, as writing large amounts of
data to slower storage devices may be costly. To reduce any unnecessary communication
and computation, Spark also included lazy evaluation, which builds the entire task
graph prior to execution to determine what needs to be computed.

% Didn't mention Kubernetes as it's still experimental
Spark is compatible with three different schedulers: Spark
Standalone, YARN and Mesos.
The Spark Standalone scheduler is a simple default scheduler built into Spark.
In contrast, the YARN scheduler is primarily designed to schedule Hadoop-based 
workflows, whereas Mesos can be used to schedule
a variety of different workflows. As researchers would likely be
executing their workflows in HPC environments with neither YARN or Mesos
installed, we limit our focus to Spark's Standalone scheduler.

The Spark Standalone scheduler is composed of three main processes: the
\emph{master}, the \emph{workers} (\emph{slaves}, in Spark) and the
\emph{driver}. The \emph{master} coordinates resources provisioned by
\emph{worker} processes on the cluster. The application is submitted to the
\emph{driver} that in turn requests workers to the master and dispatches tasks
to them. A job is divided into stages to be executed in a different process onto
the workers. Each stage's operation is represented as a high-level task in the
computation graph. The Spark standalone scheduler uses a FIFO 
(First-In-First-Out) job scheduling policy, and it has
two execution modes: client mode, where the driver runs in a dedicated process
outside of the Spark cluster, and cluster mode, where the driver runs within
a worker process. Our experiments use client mode as cluster mode is not
available in PySpark.

Python is commonly selected as the programming language of choice in scientific
communities, and in particular, for our use case of neuroscience, where
numerous specialized Python libraries exist to study the data.
While serialization of Python to Java may lead to significant overheads, we
chose to focus on PySpark API due to its suitability for neuroimaging
research.

We used Apache Spark v2.4.0.

\subsubsection{Dask} Dask is a Python-based Big Data engine that is becoming
increasingly popular in the scientific Python ecosystem. Like Spark, Dask avoids data transfers and needless computation and communication through
in-memory computing, data locality, and lazy evaluation. Dask workflows can 
further reduce data transfer costs by leveraging multithreading whenever 
communication is not bounded by Python's GIL. Dask relies on data 
lineage to achieve fault tolerance, however, unlike Spark, it does not require
operations to be coarse-grained. Furthermore, Dask is lightweight and modular,
allowing users to only install the components they require.

Dask has five main data structures:
\href{https://docs.dask.org/en/latest/array.html}{Array}, \href{https://docs.dask.org/en/latest/bag.html}{Bag},
\href{https://docs.dask.org/en/latest/dataframe.html}{DataFrame},
 \href{https://docs.dask.org/en/latest/delayed.html}{Delayed}, 
 and \href{https://docs.dask.org/en/latest/futures.html}{Futures}. A Dask
Array is used for the
processing of large arrays. It provides a distributed clone of the popular NumPy
library. Similar to Spark's RDD, a Dask
Bag is a parallel
collection of Python objects. It offers a programming abstraction similar to the
\href{https://toolz.readthedocs.io/en/latest/}{PyToolz library}. A Dask
Dataframe is a parallel
composition of
\href{http://pandas.pydata.org/pandas-docs/stable/reference/api/pandas.DataFrame.html}{Pandas Dataframes},
 used to process a large amount of tabular data.
Dask Delayed supports
arbitrary tasks
that do not fit in the Array, DataFrame or Bag APIs. Finally,
Dask Futures are similar to
Delayed in that they support arbitrary tasks, but they operate in real-time
rather than lazily. Except for Dask Bag, all APIs, by default, use
the local multithreaded scheduler. Dask Bag, instead, relies on the local
multiprocessing scheduler.  All Dask data structures, except for
Dask Array and Dask DataFrame, were used in our experiments.

The Dask graph is the internal representation of a Dask application to be
executed  by the scheduler. API operations generate multiple small tasks in
the computation graph, allowing an easier representation of complex
algorithms.

The Dask engine is compatible with numerous distributed schedulers, including 
\href{https://github.com/dask/dask-yarn}{YARN} and
\href{https://github.com/mrocklin/dask-mesos}{Mesos}, similarly to Spark. Dask 
also provides its own distributed scheduler, known as the Dask Distributed
scheduler. Although Dask is compatible with schedulers commonly found in HPC
clusters, we chose to use Dask's Distributed scheduler, to keep the environment
balanced between both engines.

In the Dask Distributed scheduler, a process called \textit{dask-scheduler}
administrates the resources provided by \textit{workers} in the cluster. The
scheduler receives jobs from clients and assigns tasks to available workers.
Similarly to Spark's scheduler, Dask Distributed completes the processing of a
graph branch before moving along to the next one.

In our experiments, we used Dask v1.1.4.

\subsection{Infrastructure}

 We used Compute Canada's
 \href{https://docs.computecanada.ca/wiki/Cloud\_resources}{Arbutus Cloud} operated by
 the \href{https://www.westgrid.ca}{WestGrid} regional organization at the University
 of Victoria, running OpenStack Queens release. We used c8-30gb-186 cloud
 instances with 8 VCPUs, an Intel Xeon Gold 6130 processor, \SI{30}{\giga\byte} of
 RAM at \SI{2666}{\mega\hertz}, \SI{20}{\giga\byte} of mounted storage, and a base
 image running CentOS 7.5.1804 with Linux kernel version
 3.10.0\-862.11.6.el7.x86\_64. Instances were connected by a
 \SI{10}{\giga\bit/\second} Ethernet network.

 Cloud instances hosted a single Dask or Spark worker, configured to use 8 CPUs. We
 used Dask's default configuration that uses all the available memory on the
 instance. Its default heuristic is to: target a 60\% memory load, spill to disk at
 70\%, pause the worker at 80\%, and terminate the worker at 95\%. We configured
 Spark to use 1 executor per worker and \SI{25}{\giga\byte} of memory per executor, to leave \SI{5}{\giga\byte}
 for off-heap. We configured the Spark driver to use \SI{25}{\giga\byte} of memory, and used the
 default configuration for the master. We used the default configuration for worker
 memory management: at 60\% it spills data to disk, and 50\% of that amount is
 reserved for storage that is immune to eviction.

 One cloud instance did not host any worker and had a \SI{2}{\tera\byte} disk volume
 shared with the other instances using the Network File System (NFS) v4.
 This instance was also used for the Spark driver and master, for the
 Dask scheduler, and for job monitoring with the Spark and Dask user
 interfaces. For both Spark and Dask, spilled data was evicted to the NFS.

\subsection{Dataset}

We used BigBrain~\cite{Amunts:13}, a three-dimensional image of a human
brain with voxel intensities ranging from 0 to 65,535. The original data is
stored in 125 blocks in the MINC~\cite{minc} HDF5-based format, available
at \url{ftp://bigbrain.loris.ca/BigBrainRelease.2015/3D\_Blocks/40um} at a
resolution of \SI{40}{\micro\metre}. We converted the blocks into the
\href{https://nifti.nimh.nih.gov/nifti-1}{NIfTI} format, a popular format
in neuroimaging. We left the NIfTI blocks uncompressed, resulting in 
a total data size of \SI{81}{\giga\byte}. 
To evaluate the effect of block size, we resplit these blocks into 30, 125 and 750 blocks of 
\SI{2.7}{\giga\byte}, \SI{0.648}{\giga\byte}, and
\SI{0.108}{\giga\byte}, using the \href{https://github.com/big-data-lab-team/sam}{sam} library.

We also used the dataset provided by the Consortium for Reliability and
Reproducibility (\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR}~\cite{zuo2014open}) as
available on \href{http://datasets.datalad.org/?dir=/corr/RawDataBIDS}{DataLad}. The
entire dataset is \SI{408.4}{\giga\byte}, containing anatomical, diffusion and
functional images of 1,397 subjects acquired in 29 sites. We used all 3,491 anatomical
images, representing \SI{39}{\giga\byte} overall (\SI{11.17}{\mega\byte} per image on
average).

\subsection{Applications}

We used three neuroimaging applications to evaluate the engines in different
conditions. The first two, incrementation and histogram, are simple 
synthetic applications representing map-only and map-reduce applications,
respectively. The third, is a real neuroimaging application representative
of popular BIDS applications~\cite{gorgolewski2017bids}. All scripts used for our
experiments are available at
\url{https://github.com/big-data-lab-team/paper-big-data-engines}

%% Incrementation
\subsubsection{Incrementation}
We used an adaptation of the image incrementation pipeline used
in~\cite{hayot2019performance} (see Algorithm~\ref{alg:incrementation}).
The application reads blocks of the BigBrain image from the shared file
system, increments the intensity value of each voxel by 1 to avoid caching
effects, sleeps for a configurable amount of time to emulate more
compute-intensive processing, repeats this process for a specified number of
iterations, and
finally writes the result as a NIfTI image back to the shared file system.
This application allows us to study the behavior of the engines when all
inputs are processed independently, in a map-only scenario (see
Figure~\ref{fig:tg-inc}). This mimics the behavior of analyzing multiple
independent subjects in parallel.

\begin{algorithm}[!b]
    \caption{Incrementation (adapted from~\cite{hayot2019performance})}\label{alg:incrementation}
    \begin{algorithmic}
    \Require{\(x\), a sleep delay in float}
    \Require{\(file\), a file containing a BigBrain block}
    \Require{\(fs\), NFS mount to write image to}
    \State{Read \(block\) from \(file\)}
    \ForEach{\(i \in iterations\)}
        \ForEach{\(block \in image\)}
            \State{\(block\gets block+1\)}
            \State{Sleep \(x\)}
        \EndFor
    \EndFor
    \State{Write \(block\) to \(fs\)}
\end{algorithmic}
\end{algorithm}

\begin{figure}[!b]
    \centering
    \includegraphics[height=\columnwidth,
    angle=-90]{images/incrementation-task-graph.png}
    \caption{Task graph for Incrementation with 5 iterations and 3 BigBrain blocks.
    Circles represent the incrementation and sleep function while rectangles
    represent data elements.}\label{fig:tg-inc}
\end{figure}

\subsubsection{Histogram}

 Our second application calculates the histogram of the BigBrain image (see
 Algorithm~\ref{alg:histogram}). It reads the image blocks from the shared
 file system, calculates intensity frequencies, aggregates the frequencies
 across the blocks, and finally writes the resulting histogram to the
 shared file system as a single \SI{766}{\kilo\byte} file. It follows the
 map-reduce paradigm, in which the final result is obtained from all the
 input blocks (see Fig.~\ref{fig:tg-histo}). This application requires data
 shuffling, incurring additional performance costs compared to
 Algorithm~\ref{alg:incrementation}. The total amount of shuffled data is,
 however, limited to \SI{2.62}{\mega\byte} per block as it only consists
 of image histograms. Two implementations are studied: a pure Python one,
 where frequencies are computed through Python dictionary manipulations, and one based
 on the \href{https://github.com/numpy}{NumPy} library, that implements array computations in C.

\begin{algorithm}[!t]
    \caption{Histogram}\label{alg:histogram}
    \begin{algorithmic}
    \Require{\(files\), files containing BigBrain blocks}
    \Require{\(fs\), NFS mount to save image to}
    \ForEach{\(file \in files\)}
        \State{Read \(block\) from \(file\)}
        \State{{Calculate \(frequencies\) in \(block\)}}
    \EndFor

    \State{\(histogram\gets\)Aggregate \(frequencies\) of each \(file\)}

    \State{Write \(histogram\) to \(fs\)}
    \end{algorithmic}
\end{algorithm}

\begin{figure}[!t]
    \centering
    \includegraphics[height=\columnwidth,
    angle=-90]{images/histogram-task-graph.png}
    \caption{Task graph for Histogram with 3 BigBrain blocks. Circles represent the
    functions while rectangles represent data elements.
    }\label{fig:tg-histo}
\end{figure}

\subsubsection{BIDS app example}

We used the
\href{https://github.com/BIDS-Apps/example}{BIDS app example} map-reduce style 
application. The map phase, also called participant analysis, extracts the brain
tissues of a subject's 3D MRI from the CoRR dataset using the FMRIB Software Library
(\href{https://fsl.fmrib.ox.ac.uk/fsl/fslwiki}{FSL}), and writes the resulting
image (\SI{2.47}{\mega\byte} on average) to the NFS. The reduce phase, or group
analysis, computes the volume of each brain and returns the average
volume, shuffling a total of \SI{8.6}{\giga\byte} image data.

While Incrementation and Histogram were implemented directly in Python, the 
BIDS App example requires an external set of command-line tools distributed as a Docker container image
(bids/base\_fsl on DockerHub). We converted the Docker image to a Singularity image for use in 
an HPC environment, using
\href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}.
The image was preloaded on the NFS and Singularity version 3.2.1\-1.el7 was installed
for all instances.

\subsection{Experiments}

Four parameters were varied in our experiments, as shown in
Table~\ref{tab:param}. We varied (1) the number of workers, to assess the
scalability of the engines, (2) the number of BigBrain blocks in
Incrementation and Histogram, to measure the effect of different I/O patterns
and parallelization degrees, and (3) the number of iterations and sleep
delay in Incrementation, to evaluate the effect of job length and number of
tasks.
It should be noted that increasing the number of blocks or iterations also
increases the total compute time of the application for a given sleep
delay. To avoid any potential external bias such as background load on the
network, we ran the experiments in a randomized order and cleared the page
cache of each worker before each execution.

For each run, we measured the application makespan as well as the cumulative 
data read, compute, data write, and engine overhead times across all application
tasks. 
The overhead calculation for each CPU is the end time of its last processed task
minus the total runtime of the tasks it executed. The summation of those results is the
total overhead.

% The baseline for our experiments is: 8 workers, 125 blocks, 10 iterations and 4
% seconds sleep delay.

\begin{table}[!t]
    \renewcommand{\arraystretch}{1.3}
    \caption{Parameters for the experiments}\label{tab:param}
    \centering
    \begin{tabular*}{\columnwidth}{llll}
    \hline
                        & Incrementation & Histogram             & BIDS Example          \\ \hline
    \# of worker        & 1, 2, 4, 8     & 1, 2, 4, 8            & 1, 2, 4, 8            \\
    \# of blocks        & 30, 125, 750   & 30, 125, 750          & \multicolumn{1}{c}{-} \\
    \# of iterations    & 1, 10, 100     & \multicolumn{1}{c}{-} & \multicolumn{1}{c}{-} \\
    Sleep delay {[}\SI{}{\second}{]} & 1, 4, 16, 64   & \multicolumn{1}{c}{-} & \multicolumn{1}{c}{-} \\ \hline
    \end{tabular*}
    \end{table}

%%%%% RESULTS %%%%%
\section{Results}

%%% INCREMENTATION %%%
%% worker
\subsection{Incrementation: Number of workers}
Figure~\ref{fig:inc_ms_worker} shows the makespan of the Incrementation application
for different numbers of workers and engines. The bars show the average
makespan over 3 repetitions while the error bars are the standard deviations. Overall,
there is no substantial makespan difference between the engines. Dask seems to have a
slight advantage over Spark, \SI{83.61}{\second} on average,
with Delayed and Futures being slightly better than Bag.

For all engines, the makespan is far from decreasing linearly with the
number of workers. The makespan
even increases between 4 and 8 workers. This is due to the high impact of
data transfers and engine overhead on the application.

Figure~\ref{fig:inc_tt_worker} shows the total execution time of the
Incrementation application, broken down into data transfer (read and
write), compute (sleep), and overhead time. As expected, the computing time
stays similar when the number of workers increases. However, the data
transfer time and overhead increase proportionally to the number of
workers with regression slopes: Spark (\SI{2337}{\second/task}), Bag
(\SI{2650}{\second/task}), Delayed (\SI{3251}{\second/task}), Futures
(\SI{3570}{\second/task}).

On Figure~\ref{fig:inc_tt_worker}, we also note that Spark's overhead is
slightly lower than Dask's, particularly when the number of workers
increases. Moreover, Delayed and Futures have a higher overhead than Bag.
However, overhead differences are compensated by an increase in data
transfer time, as a reduced overhead increases the concurrency between data
transfers and the application saturates the disk bandwidth of the shared file system.

\begin{figure}[!b]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_worker.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_worker}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_worker.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_worker}
    \end{subfigure}
    \caption{125 blocks, 10 iterations, \SI{4}{\second} delay, 8
    CPUs/worker}\label{fig:inc_worker}
\end{figure}

% blocks
\subsection{Incrementation: Number of blocks}

Figure~\ref{fig:inc_ms_block} shows the Incrementation makespan when varying the
number of image blocks for constant BigBrain image size. We were not able to run
Spark for 30 blocks due to its \SI{2}{\giga\byte} limitation in the task size.
 Once again, we do not observe any substantial difference among the engines.
For all engines, makespan variability increases with the number of blocks, however,
engines scale very well in general.

In Figure~\ref{fig:inc_tt_block}, the total execution time of each function is shown.
For 30 blocks, the increased overhead of Delayed and Futures is a calculation artifact 
coming from the fact that only 30 of the 64 available threads can be used concurrently. Once again, the data transfer time reduces with more blocks but 
the overhead time increases by a similar amount. This is not observed for 30 blocks
as the workers are not used at full capacity, i.e., some threads are idle. Finally,
the variability of the overhead increases with the number of blocks, which explains
the makespan variability mentioned previously.

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_block.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_block}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_block.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_block}
    \end{subfigure}
    \caption{10 iterations, \SI{4}{\second} delay, 8 workers, 8
    CPUs/worker}\label{fig:inc_block}
\end{figure}

%% Iterations
\subsection{Incrementation: Number of iterations}
Figure~\ref{fig:inc_ms_itr} shows the makespan of the application while
varying the number of iterations. Overall, Spark and Dask APIs are once
again equivalent, although Delayed and Futures are slightly faster
than Bag and Spark for 1 and 10 iterations, and Futures are faster than
Delayed, Bag and Spark for 100 iterations. Differences remain minor
though.

In Figure~\ref{fig:inc_tt_itr}, the total execution time breakdown is
shown. We observe the good scalability of all the engines with the number of
iterations.

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_itr.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_itr}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_itr.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_itr}
    \end{subfigure}
    \caption{125 blocks, \SI{4}{\second} delay, 8 workers, 8 CPUs/worker}
\end{figure}

%% Sleep time
\subsection{Incrementation: Sleep delay}
Figure~\ref{fig:inc_ms_sleep} shows the makespan of the Incrementation application
for different sleep delays. Overall, all engines again perform the same and scale
well with task duration. Spark is initially slower than the Dask APIs, however, it is
faster with an increased sleep delay. Also, within the Dask APIs, Dask Bag is
slower than the other two, but it is not considerable.

Figure~\ref{fig:inc_tt_sleep} shows the total execution time breakdown. Spark has the
smallest overhead. As previously observed, variations in overhead time are almost
exactly compensated by variations in data transfer time.

\begin{figure}[!b]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_sleep.png}%
        \caption{Incrementation makespan}\label{fig:inc_ms_sleep}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/inc_idle_sleep.png}%
        \caption{Incrementation total time}\label{fig:inc_tt_sleep}
    \end{subfigure}
    \caption{125 blocks, 10 iterations, 8 workers, 8 CPUs/worker}
\end{figure}

%% Baseline gantt chart
\subsection{Incrementation: Gantt chart}

Figure~\ref{fig:inc_gantt} shows the Gantt chart obtained for each engine. Gantt
charts are structured in batches of up to 64 read-compute-write concurrent sequences.
File reads in the first batch are much longer than the ones in the following batches:
this is due to the high synchronization of data transfers that leads to a high
saturation of the shared file system. We also note that overhead, represented in
white, is concentrated around the data transfer tasks and the
computing tasks that run concurrently with data transfers.

\begin{figure}[!b]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \href{https://mathdugre.github.io/paper-big-data-engines/spark-inc-baseline.html}{
            \includegraphics[clip,width=\columnwidth,
            height=0.15\textheight]{images/spark_inc_gantt.png}}
        \caption{Spark}\label{fig:inc_spark_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \href{https://mathdugre.github.io/paper-big-data-engines/dask-bag-inc-baseline.html}{
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/bag_inc_gantt.png}}
        \caption{Dask Bag}\label{fig:inc_dask_bag_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \href{https://mathdugre.github.io/paper-big-data-engines/dask-delayed-inc-baseline.html}{
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/delayed_inc_gantt.png}}
        \caption{Dask Delayed}\label{fig:inc_dask_delayed_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \href{https://mathdugre.github.io/paper-big-data-engines/dask-futures-inc-baseline.html}{
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/futures_inc_gantt.png}}
        \caption{Dask Futures}\label{fig:inc_dask_futures_gantt}
    \end{subfigure}
    \caption{Incrementation Gantt chart: 125 blocks, 10 iterations, \SI{4}{\second} delay, 8 workers,
8 CPUs/worker. All Gantt charts are clickable and link to interactive figures with additional information.}\label{fig:inc_gantt}
\end{figure}

%%% HISTOGRAM %%%
\subsection{Histogram (Python): Number of workers}
Figure~\ref{fig:histo_ms_worker} shows the makespan of the Histogram application for various
amounts of workers. Spark is significantly faster than Dask APIs. The difference
narrows as the number of workers increases but it remains significant. Between Dask APIs
there is no substantial difference, however, Bag is slightly faster on average.

From Figure~\ref{fig:histo_tt_worker}, the total time spent in each function is
shown. The computing time is significantly larger for Dask than for Spark. This is
presumably due to Python's GIL preventing Dask to parallelize the computation
over
multiple threads. Overall, the I/O and overhead times are comparable for all engines.

Figure~\ref{fig:histo_tt_worker} shows that Dask engines have a similar
total execution time when their number of workers varies, except for 8
workers where it increases slightly due to increased read time.
On the other hand, Spark total execution time keeps increasing; especially at 8
workers. This is because Dask engines benefit more from additional workers, in
this GIL-bounded scenario.

>>>>

\title{Impact of engines scheduler on the performance of Big Data pipelines in HPC}

@@

\begin{abstract}
	% TODO
\end{abstract}

\begin{IEEEkeywords}
	Performance, Scheduler, Big Data, Dask, Spark, Ray, Neuroimaging
\end{IEEEkeywords}
\MD{Discuss which keyword to use for the paper.}

\section{Introduction}

\section{Background}
\subsection{Lustre} % TODO

\subsection{Dask}
Dask is a Python-based Big Data engine with growing popularity in the scientific Python ecosystem.
To mitigate the data transfer bottleneck in Big Data workflows, Dask was designed with data locality and in-memory in mind.
Data locality, popularized by Map-Reduce \cite{dean2008mapreduce}, schedules tasks where the data reside.
In-memory computing minimize the overhead of transfering data to disk by keeping data in memory when possible.
Dask uses lazy evalutation to reduce unnecessary communication and computation.
The engine builds a dynamic graph prior to execution allowing to determine which task to compute.
Dask workflows can further reduce data transfer by using  leveraging mulithreading whenever the application is not bounded by Python's GIL.
Fault-tolerance is achieved by recording of data lineage, the sequence of operations used on to modify the initial data.

Dask offers five data structures: Array, Bag, DataFrame, Delayed, and Futures. \MD{Insert link to the API.}
The Dask Array offers a clone of NumPy API for distributed processing of large arrays.
A Dask Bag is a distributed collection of Python object that offers a programming abstraction similar to PyToolz. \MD{Link to PyToolz.}
Dask Dataframe is a parallel composition of Pandas Dataframes used to process large amount of tabular data.
Dask Delayed offers an API for distributing abrirtary function that do not fit in the above frameworks.
Lastly, Dask Futures can also execute abrirtary functions, however, it launch computation immediately rather than lazily.
Dask modularity allow users to install only required components making it lightweigth.

In Dask a scheduler decides where and when to execute tasks using the Dask graph.
API operations generate multiple fine-coarsed tasks in the computation graph allwing an easier representatin of complex algorithms.

The Dask engine is compatible to numerous distributed schedulers, including YARN and Mesos.
Dask also provide its own \textit{Dask Distributed scheduler}.
Althought Dask is compatible with schedulers available in HPC clusters,
We chose to use Dask Distributed scheduler, to keep the environment balanced between the engines.

In Dask Distributed scheduler, a \textit{dask-scheduler} process admistrates the resource provided by \textit{dask-worker} in the cluster.
The scheduler receives jobs from clients and assigns tasks to available workers.
Dask scheduler use a FIFO (First-In-First-Out) job scheduling policy.
That is, completely process a branch of the Dask graph before proceding with the next one.

Dask offers multiple ways to deploy a cluster including: with SSH configs, Kubernetes, SLURM, PBS, and others.
For our experiments, we used the \href{https://jobqueue.dask.org/en/latest/generated/dask_jobqueue.SLURMCluster.html}{Dask SLURM cluster} API.

Dask \MD{Add version when after locking it for experiments.} was used.

\subsection{Ray}

\subsection{Apache Spark}
Apache Spark is a widely-used general pupose Big Data engine.
Like Dask, it aims at reducing data transfer cost by incorporating data locality and in-memory computing, and lazy evaluation.

Spark primary data structure is Resilient Distributed Dataset (RDD)\cite{RDD}, a fault-tolerant, parallel collection of data elements.
RDDs is the basis of the other Spark data structure: Datasets and DataFrames.
Datasets are similar to RDD but benefits from additional performance by leveraging the Spark SQL's optimized execution engine.
The DataFrame API are Datasets organized into named-columns and used to process tabular data.
While the DataFrame API is available in all supported language, DAtasets are limited to Sacla and Java.

Python is a common programming language in the scientific community.
This is particulary true in the neuroimaging field with numerous specialized Python libraries to operate on data.
While serialization from Python to Java create overhead, we found it to be minimal. \MD{cite SC19 paper.}
We chose to focus on PySpark API to have a more balance environment between the different engines and due to its suitability to neuroimaging.

\section{Methods}
\subsection{Infrastructure} % TODO
\subsection{Dataset} % TODO

\subsection{Applications}
\subsubsection{Increment}
We adapted the increment pipeline used in \cite{hayot2019performance}.
This synthetic application read blocks of the BigBrain from Lustre and sleeps for specified period.
To simulate intermediate results, we repeat the sleep process for a configurable amount of time.
We prevent data caching of the blocks by incrementing their voxels value by 1 after each sleep operation.
Finally, we write the resulting NIfTI image back to Lustre.
This application allow us the study engines when the inputs are processed independently.
The map-only scenario of this application mimics the processing of multiple independt subjects in parralel.
\MD{TODO: Add the Figure for the application graph.}

\subsubsection{Multi-Increment}
Our second application is a revision of the increment pipeline.
A major difference is that, at each iteration, a random BigBrain blocks is used to as the increment value.
This change allows the multi-increment application to have inter-worker communication while remaining simplistic.
\MD{TODO: Add the Figure for the application graph.}

\subsubsection{Histogram}
Our third application calculatest the histogram of the BigBrain image.
It reads the BigBrain blocks from Lustre, calculate frequency of each intensity, then write the aggregated result back on Lustre.
This map-reduce application has a very high read over write ratio.
Moreover, this application requires data shuffling; although little.
This place it in between the increment and multi-increment pipeline in terms of inter-worker communication.
\MD{TODO: Add the Figure for the application graph.}

\subsubsection{Kmeans} % TODO
\MD{TODO: Write about it once implemented.}

\subsubsection{BIDS app example}
Our last application is BIDS App example: a neuroimaging pipeline to extract brain volume.
For this application, the CoRR dataset was used.
It extracts the brain volume of each participant, then computes the average for each group of participants.
Unlike the other application, BIDS App example is a set of command-line distributed in a Docker image (bids/example on DockerHub).
We converted the Docker image to a Singularity image for use in HPC environment, \MD{Cite paper on reason why this is done.}
using \href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}
\MD{Need to discuss if we preload the image on the worker nodes.}
\MD{TODO: Add the Figure for the application graph.}

\subsection{Experiments} % TODO

\section{Results} % TODO

\section{Discussion} % TODO

\section{Conclusion} % TODO

\section*{Acknowledgment} % TODO
